---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

###Load the data

```{r echo=TRUE}
library(data.table)
library(ggplot2)
library(knitr)

setwd("C:/Users/Carolina/Documents/Data scientist/reproductible research/RepData_PeerAssessment1-master/RepData_PeerAssessment1-master/")
activityData <- read.csv("activity.csv")
```


###Process/transform the data into a format suitable for your analysis

```{r echo=TRUE}
activityData <- data.table::fread(input = "activity.csv")
```


## What is mean total number of steps taken per day?



```{r echo=TRUE}
library(ggplot2)

total.steps <- tapply(activityData$steps, activityData$date, FUN = sum, na.rm = TRUE)
qplot(total.steps, binwidth = 1000, xlab = "total number of steps taken each day")
```

```{r echo=TRUE}
mean(total.steps, na.rm = TRUE)
```

```{r echo=TRUE}
median(total.steps, na.rm = TRUE)
```



## What is the average daily activity pattern?



```{r echo=TRUE}
library(ggplot2)
averages <- aggregate(x = list(steps = activityData$steps), by = list(interval = activityData$interval), 
    FUN = mean, na.rm = TRUE)
ggplot(data = averages, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + 
    ylab("average number of steps taken")
```

On average, which 5-minute interval contains the maximum number of steps?

```{r echo=TRUE}
averages[which.max(averages$steps), ]
```


## Imputing missing values


```{r echo=TRUE}
missing <- is.na(activityData$steps)
# How many missing
table(missing)
```

The missing values are filled in with mean value for the 5-minute interval.


```{r echo=TRUE}
# Replace each missing value with the mean value of its 5-minute interval
fillvalue <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps)) 
        filled <- c(steps) else filled <- (averages[averages$interval == interval, "steps"])
    return(filled)
}
filledData <- activityData
filledData$steps <- mapply(fillvalue, filledData$steps, filledData$interval)
```

Using the filled data set, we are making an histogram of the total number of steps  each day and calculate the mean and median.


```{r echo=TRUE}
totalSteps <- tapply(filledData$steps, filledData$date, FUN = sum)
qplot(totalSteps, binwidth = 1000, xlab = "total number of steps taken each day")
```

```{r echo=TRUE}
mean(totalSteps)
```

```{r echo=TRUE}
median(totalSteps)
```



## Are there differences in activity patterns between weekdays and weekends?


```{r echo=TRUE}
filledData$day=ifelse(as.POSIXlt(as.Date(filledData$date))$wday%%6==0,
                          "weekend","weekday")
# For Sunday and Saturday : weekend, Other days : weekday 
filledData$day=factor(filledData$day,levels=c("weekday","weekend"))
```


```{r echo=TRUE}
stepsInterval2=aggregate(steps~interval+day,filledData,mean)
library(lattice)
xyplot(steps~interval|factor(day),data=stepsInterval2,aspect=1/2,type="l")

```
